---
title: "Creating New ARDs"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The basic functions offered by {cards} can create a wide variety of ARDs.
However, sometimes we may need to include the outputs from more complicated statistical methods in our ARDs.
In this article we'll look at a few different ways to implement the output from these statistical methods.

## {cardx}

The {cardx} package is an extension of {cards}.
The idea is that {cards} provides the core functions to create ARDs, while {cardx} contains a large number of extensions that implement various, commonly used statistical methods.
There are a large number of extensions for a wide variety of methods, including (but not limited to):

* Regression models
* ANOVA
* Chi-squared Test
* t-test
* LS Mean Difference
* Survival Estimates and Differences


When looking to include the output from a statistical method your first port of call should be to see if it has already been implemented in {cardx}.
You can find the full list of available functions [here](https://insightsengineering.github.io/cardx/main/reference/index.html).


Consider a simple t-test comparing the mean age (`AGE`) across two treatments arms (`ARM`).
In {cardx} we have the function `cardx::ard_stats_t_test()`

```{r cardx}
cards::ADSL |>
  dplyr::filter(ARM %in% c("Xanomeline High Dose", "Xanomeline Low Dose")) |>
  cardx::ard_stats_t_test(by = ARM, variables = AGE)
```

In the output, we see the outputs from the t-test; the mean difference, confidence interval limits and p-value.
It's also useful to see the functions inputs; for example, we can see that we did not use equal variances as the `stat` is `FALSE` for `stat_name` `var.equal`.
This is useful for re-use, if we need to run the test again we can use the ARD to see what options we need to use to recreate the result. 

# Create a new `ard_*()` function

_But what do we do if the statistical method that we want to use hasn't been implemented already in {cardx}?_

Implementing a new function to create an ARD for a statistical method is often simple; all we need to do is write a function that outputs the results as a named list! 

We'll first look at how `broom::tidy()` can make this even easier for us, then provide an example on how to implement from scratch.

## Using `broom::tidy()`

Typically, a user will pass a function which returns a scalar value in the `cards::ard_continuous(statistic)` argument.
However, the argument also allows for functions that return named lists, where the names from the list will then be used as the statistic names.
Since data frames or tibbles are just named lists with a little more formatting, we can pass a function to the `statistic` argument which returns a single row data frame or tibble and the behavior will be the same.
Each column of the table gives the name of the statistic via its column name, and then the corresponding value. 

Commonly used statistical methods outputs are able to be passed through `broom::tidy()`, which will convert the output into a tibble.
We can then pass the output of `broom::tidy()` through to the `cards::ard_continuous(statistic)` argument of the ARD function we wish to use, this leads to an ARD output like we see in the above example, where we have one row per relevant input or output from the statistical method.

Please note that we can only use the `broom::tidy()` output directly, as we see in the below example, when the output is a tibble with a single row.

Let's extend our t-test example from above.
This time we want to carry out a one-sample t-test.
We can just pass the code to carry out the one-sample t-test and pass the output through `broom::tidy()` to the `statistic` argument like so:

```{r tidy-stats-method}
cards::ADSL |>
  dplyr::filter(ARM %in% c("Xanomeline High Dose", "Xanomeline Low Dose")) |>
  cards::ard_continuous(
    variables = AGE,
    statistic = everything() ~ list(t_test = \(x) t.test(x) |> broom::tidy())
  ) |>
  dplyr::mutate(context = "t_test_one_sample")
```

In the above chunk of code, if we focus on what we pass to the `statistic` argument:

* `everything()` means we want to run this test on all columns passed to the `variables` argument.
* `t.test(x)` is the function which carries out the statistical test, in this case with the default arguments.
* We pipe the output of `t.test(x)` into `broom::tidy()` which converts the output of the test into a tibble (which, remember, is just a named list!)
* The function we've defined (`\(x) t.test(x) |> broom::tidy()`) itself needs to be included in a named list, where here we've chosen to use the label `'t_test'`.

Over 100 different statistical methods implemented in R are able to be 'tidied' using `broom::tidy()`.
However, the method you aim to use might not be, or the current `broom::tidy()` implementation might not contain the information that you need to be in your ARD.
In that case we'll have to format the output ourselves.

## Without `broom::tidy()`

As mentioned above, we need to define a function which carries out or required statistical method and outputs a named list of the information we wish to include in the ARD.

As an example, let's write a function which carries out a Wilcoxon signed rank test over one variable using the function `wilcox.test`.
As an output we just want to record the method and the p-value.

```{r wilcox-function}
wilcox_one_var <- \(x) wilcox.test(x)[c("method", "p.value")]
```

Let's now use this function when creating an ARD with {cards}.
Remember we just need the statistic to be a named list, so we'll call our function inside a named list.
We also don't need to specify any arguments, in this case it will pick up that the one variable `x` corresponds to the data we are testing, in this case `AGE` for the individual treatment arms.

```{r wilcox-with-cards}
cards::ADSL |>
  cards::ard_continuous(
    variables = AGE,
    by = ARM,
    statistic = ~ list(wilcox = wilcox_one_var)
  )
```

We see here that we get an output of 8 rows, 2 rows (one for the method, and one for the p-value) for each of the 4 treatment arms.

## Complex inputs

The examples above are great to illustrate a simple case, but it is perhaps a rare scenario where we are implementing a statistical method with a single vector as its input.
How would we update the code above if we need to implement a two-sample t-test?

The `cards::ard_complex()` is similar to `cards::ard_continuous()`, but allows for more complex inputs in the function passed in the `statistic` argument.
In `cards::ard_continuous()`, the functions passed must accept a single vector, e.g. `\(x) t.test(x)`.
But in `cards::ard_complex()`, in addition to the vector being passed, the `data` subset, the `full_data`, character `by`, and character `strata` are also passed (see the `cards::ard_complex()` for a full description). Your function does not need to utilize each of these elements, but each _will be passed_ to your function.
As a result, we recommend your function accept the triple dots to handle unused arguments.

An implementation of a two-sample t-test may look like this:

```{r}
ttest_two_sample <- \(x, data, ...) t.test(x ~ data[["ARM"]]) |> broom::tidy()

cards::ADSL |>
  dplyr::filter(ARM %in% c("Xanomeline High Dose", "Xanomeline Low Dose")) |>
  cards::ard_complex(
    variables = AGE,
    statistic = everything() ~ list(t_test = ttest_two_sample)
  ) |>
  dplyr::mutate(context = "t_test_two_sample")
```

# Handling errors

Let's consider what happens when we encounter an error in our statistical method.

```{r wilcox-fn-error}
wilcox_one_var_error <- function(x) {
  stop("AN ERROR!")
  wilcox.test(x)[c("method", "p.value")]
}

cards::ADSL |>
  cards::ard_continuous(
    variables = AGE,
    by = ARM,
    statistic = ~ list(wilcox = wilcox_one_var_error)
  )
```

In the output we see that we only get 4 rows of output, the error has been stored in the `error` column but `stat_name` and `stat_label` now just take the list name of "wilcox" that we define in the statistic argument.
This could have unintended effects in downstream code, we may be relying on the `stat_name` and `stat_label` having values of "method" and "p-value", or just that the output has 2 rows per treatment arm.

To handle this we can specify the expected results from our function, so that even if we encounter an error during the code run we can be assured that the output will be of a consistent format so as not to impact downstream code.

Here's an example of how to specify the expected output using `cards::as_cards_fn()`:

```{r as_cards_fn}
wilcox_one_var_error <- cards::as_cards_fn(
  wilcox_one_var_error,
  stat_names = c("method", "p.value")
)

cards::ADSL |>
  cards::ard_continuous(
    variables = AGE,
    by = ARM,
    statistic = ~ list(wilcox = wilcox_one_var_error)
  )
```

Our function becomes the first argument to `cards::as_cards_fn()`, then the second argument is `stat_names` where we specify the expected names of the output list.

In the output shown here, the `error` column is still populated with the error.
However, now we have the expected 8 rows and we can see that the `stat_name` and `stat_label` match the values specified in the `stat_names` argument in the `as_cards_fn()`---helping us to avoid problems in code that relies on this output.
