---
title: "Getting Started"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Suppose you need to create the table below, and need an ARD representation of the results to get started.
Here, we will review an examples for creating a basic demographics table.

To get started, load the **{cards}** package.

```{r setup}
library(cards)
```

### Demographics

```{r, echo=FALSE, message=FALSE}
gtsummary::theme_gtsummary_compact()
ADSL |>
  dplyr::select(ARM, AGE, AGEGR1, SEX) |>
  dplyr::mutate(
    ARM = factor(ARM, levels = c("Placebo", "Xanomeline Low Dose", "Xanomeline High Dose")),
    AGEGR1 = factor(AGEGR1, levels = c("<65", "65-80", ">80"), labels = c("<65", "65-80", ">80"))
  ) |>
  gtsummary::tbl_summary(
    by = ARM,
    type = AGE ~ "continuous2",
    statistic = AGE ~ c("{median} ({p25}, {p75})", "{mean} ({sd})", "{min} - {max}"),
    value = SEX ~ "F",
    label = list(SEX = "Female", AGEGR1 = "Age Group"),
    digits = ~list(p = cards::label_round(digits = 1, scale = 100, width = 4))
  ) |>
  gtsummary::add_stat_label() |>
  gtsummary::bold_labels() |>
  gtsummary::modify_header(gtsummary::all_stat_cols() ~ "**{level}**  \nN = {n}") |>
  gtsummary::modify_column_alignment(columns = -label, "right") |>
  gtsummary::as_gt() |>
  gt::cols_width(gtsummary::all_stat_cols() ~ gt::pct(23)) |>
  gt::opt_table_font(stack = "monospace-code")
```

The table above has three types of data summaries: a **continuous** variable summary for `AGE`, a **categorical** variable summary for `AGEGR1`, and a **dichotomous** variable summary for `SEX`.

#### Continuous Summaries

To get a continuous variable summary, we will use the `ard_continuous()` function from the **{cards}** package.

```{r}
df_continuous_ard <-
  ard_continuous(
    ADSL,
    by = ARM,
    variables = AGE,
    statistic = ~ continuous_summary_fns(c("median", "p25", "p75", "mean", "sd", "min", "max"))
  )
df_continuous_ard |> head(5)
```

#### Categorical Summaries

To get the categorical variable summary, we will use the `ard_categorical()` function.

```{r}
df_categorical_ard <-
  ard_categorical(
    ADSL,
    by = ARM,
    variables = AGEGR1
  )
df_categorical_ard |> head(5)
```

#### Dichotomous Summaries

To get the dichotomous variable summary, we will use `ard_dichotomous()`.
In this case, we want to show the Female (`"F"`) level of the `SEX` variable and specify this with the `values` argument.

```{r}
df_dichotomous_ard <-
  ard_dichotomous(
    ADSL,
    by = ARM,
    variables = SEX,
    value = list(SEX = "F")
  )
df_dichotomous_ard |> head(5)
```

#### Combine Results

As a last step, you can combine all of these objects into a single object using `bind_ard()`, which is similar to `dplyr::bind_rows()` and includes additional structural checks for our results.

```{r}
bind_ard(
  df_continuous_ard,
  df_categorical_ard,
  df_dichotomous_ard
)
```

#### Shortcut

The `ard_stack()` function provides a shortcut to perform the calculations above in a single step.

In the example below, the `data` and `.by` arguments are passed to each subsequent `ard_*()` function call.
Moreover, this will also return the univariate tabulation of the `.by` variable, which would be used to add counts to the header row of the table.

```{r}
ard_stack(
  data = ADSL,
  .by = ARM,
  ard_continuous(
    variables = AGE,
    statistic = ~ continuous_summary_fns(c("median", "p25", "p75", "mean", "sd", "min", "max"))
  ),
  ard_categorical(variables = AGEGR1),
  ard_dichotomous(variables = SEX, value = list(SEX = "F"))
)
```

### Adverse Events

Next, we will review several examples for creating basic adverse events (AE) tables.
We will skip to examples utilizing the shortcut functions `ard_stack_hierarchical()` and `ard_stack_hierarchical_count()`.
These functions utilize multiple calls to `ard_hierarchical()` and `ard_hierarchical_count()` to calculate the needed summary statistics. 

For the computations below, we will not only make use of a subset of the `ADAE` dataset.
We will also rely on `ADSL` for the full study population, which is used as the denominator in the rate calculations.

To match the treatment arm variables, we need to do a small data manipulation on the naming of the treatment variable. 

```{r}
# rename trt variable
adsl <- ADSL |>
  dplyr::rename(TRTA = TRT01A)

# subset to Treatment emergent AES
adae <- ADAE |>
  # keep the most reported AEs for a smaller table
  dplyr::filter(.by = AETERM, dplyr::n() > 25, TRTEMFL == "Y") 
```

#### Participant-level summaries

A common type of AE table contains participant-level summaries.  Here, we are reporting the number and percentage of subjects with at least one AE by system organ class and preferred term.

```{r, echo=FALSE, message=FALSE}
ADAE |> 
  dplyr::filter(.by = AETERM, dplyr::n() > 25, TRTEMFL == "Y") |> 
  gtsummary::tbl_hierarchical(
    by = TRTA,
    variables = c(AEBODSYS, AEDECOD),
    id = USUBJID,
    denominator = ADSL |> dplyr::rename(TRTA = TRT01A),
    overall_row = TRUE,
    digits = ~list(p = cards::label_round(digits = 1, scale = 100, width = 4)),
    label = list(..ard_hierarchical_overall.. = "Any Adverse Event")
  ) |>
  gtsummary::modify_column_alignment(columns = gtsummary::all_stat_cols(), "right") |> 
  gtsummary::as_gt() |> 
  gt::cols_width(gtsummary::all_stat_cols() ~ gt::pct(23)) |>
  gt::opt_table_font(stack = "monospace-code")
```

The `ard_stack_hierarchical()` function provides a shortcut to perform the calculations needed for the summary table in a single step.

In the example below, the `data` and `by` arguments are passed to each subsequent calculation.
The function utilizes `USUBJID` (passed to the `id` argument) as the subject identifier for participant-level calculations, and `ADSL` is used to define the denominator.
With `variables = c(AEBODSYS, AEDECOD)`, the function returns rates of adverse events by `AEDECOD` nested within `AESOC` as well as by `AESOC`.
With `over_variables = TRUE`, the function also returns rates of any adverse event across all system organ classes and preferred terms. 

```{r}
ard_stack_hierarchical(
  data = adae,
  by = TRTA,
  variables = c(AEBODSYS, AEDECOD),
  denominator = adsl,
  id = USUBJID,
  over_variables = TRUE
)
```

#### Event-level Summaries

In addition to participant-level summaries, event-level summaries are often needed. For these types of tables, we report total counts of AEs , and therefore we can use the `ADAE` data directly.
We will need to count AEs overall, by system organ class, and by preferred term (within system organ class).


```{r, echo=FALSE, message=FALSE}
ADAE |> 
  dplyr::filter(.by = AETERM, dplyr::n() > 25, TRTEMFL == "Y") |> 
  gtsummary::tbl_hierarchical_count(
    by = TRTA,
    variables = c(AEBODSYS, AEDECOD),
    denominator = ADSL |> dplyr::rename(TRTA = TRT01A),
    overall_row = TRUE,
    digits = ~list(p = cards::label_round(digits = 1, scale = 100, width = 4)),
    label = list(..ard_hierarchical_overall.. = "Total Number of Adverse Events")
  ) |>
  gtsummary::modify_column_alignment(columns = gtsummary::all_stat_cols(), "right") |> 
  gtsummary::as_gt() |> 
  gt::cols_width(gtsummary::all_stat_cols() ~ gt::pct(23)) |>
  gt::opt_table_font(stack = "monospace-code")
```

The `ard_stack_hierarchical_count()` function provides a shortcut to perform the calculations needed for the summary table in a single step.

In the example below, the `data` and `by` arguments are passed to each subsequent calculation.
With `variables = c(AEBODSYS, AEDECOD)`, the function returns counts of adverse events by `AEDECOD` nested within `AESOC` as well as by `AESOC`.
With `over_variables = TRUE`, the function also returns counts of any adverse event across all system organ classes and preferred terms. 

```{r}
ard_stack_hierarchical_count(
  data = adae,
  by = TRTA,
  variables = c(AEBODSYS, AETERM),
  over_variables = TRUE
)
```
