---
title: "Getting Started"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Suppose you need to create the table below, and need an ARD representation of the results to get started.
Here, we will review an examples for creating a basic demographics table.

To get started, load the **{cards}** package.

```{r setup}
library(cards)
```

### Demographics

```{r, echo=FALSE, message=FALSE}
gtsummary::theme_gtsummary_compact()
ADSL |>
  dplyr::select(ARM, AGE, AGEGR1, SEX) |>
  dplyr::mutate(
    ARM = factor(ARM, levels = c("Placebo", "Xanomeline Low Dose", "Xanomeline High Dose")),
    AGEGR1 = factor(AGEGR1, levels = c("<65", "65-80", ">80"), labels = c("&lt;65", "65-80", ">80"))
  ) |>
  gtsummary::tbl_summary(
    by = ARM,
    type = AGE ~ "continuous2",
    statistic = AGE ~ c("{median} ({p25}, {p75})", "{mean} ({sd})", "{min} - {max}"),
    value = SEX ~ "F",
    label = list(SEX = "Female", AGEGR1 = "Age Group"),
    digits = gtsummary::all_categorical() ~ 0L
  ) |>
  gtsummary::add_stat_label() |>
  gtsummary::bold_labels() |>
  gtsummary::modify_header(gtsummary::all_stat_cols() ~ "**{level}**  \nN = {n}") |>
  gtsummary::modify_column_alignment(columns = -label, "right") |>
  gtsummary::as_gt() |>
  gt::cols_width(gtsummary::all_stat_cols() ~ gt::pct(23)) |>
  gt::opt_table_font(stack = "monospace-code")
```

The table above has three types of data summaries: a **continuous** variable summary for `AGE`, a **categorical** variable summary for `AGEGR1`, and a **dichotomous** variable summary for `SEX`.

#### Continuous Summaries

To get a continuous variable summary, we will use the `ard_continuous()` function from the **{cards}** package.

```{r}
df_continuous_ard <-
  ard_continuous(
    ADSL,
    by = ARM,
    variables = AGE,
    statistic = ~ continuous_summary_fns(c("median", "p25", "p75", "mean", "sd", "min", "max"))
  )
df_continuous_ard |> head(5)
```

#### Categorical Summaries

To get the categorical variable summary, we will use the `ard_categorical()` function.

```{r}
df_categorical_ard <-
  ard_categorical(
    ADSL,
    by = ARM,
    variables = AGEGR1
  )
df_categorical_ard |> head(5)
```

#### Dichotomous Summaries

To get the dichotomous variable summary, we will use `ard_dichotomous()`.
In this case, we want to show the Female (`"F"`) level of the `SEX` variable and specify this with the `values` argument.

```{r}
df_dichotomous_ard <-
  ard_dichotomous(
    ADSL,
    by = ARM,
    variables = SEX,
    value = list(SEX = "F")
  )
df_dichotomous_ard |> head(5)
```

#### Combine Results

As a last step, you can combine all of these objects into a single object using `bind_ard()`, which is similar to `dplyr::bind_rows()` and includes additional structural checks for our results.

```{r}
bind_ard(
  df_continuous_ard,
  df_categorical_ard,
  df_dichotomous_ard
)
```

#### Shortcut

The `ard_stack()` function provides a shortcut to perform the calculations above in a single step.

In the example below, the `data` and `by` arguments are passed to each subsequent `ard_*()` function call.
Moreover, we will also be returned the univariate tabulation of the `by` variable, which would be used to add counts to the header row of the table.

```{r}
ard_stack(
  data = ADSL,
  .by = ARM,
  ard_continuous(
    variables = AGE,
    statistic = ~ continuous_summary_fns(c("median", "p25", "p75", "mean", "sd", "min", "max"))
  ),
  ard_categorical(variables = AGEGR1),
  ard_dichotomous(variables = SEX, value = list(SEX = "F"))
)
```

### Adverse Events

Next, we will review several examples for creating basic adverse events (AE) tables.

#### Participant-level summaries

A common type of AE table contains participant-level summaries.  Here, we are reporting the number and percentage of subjects with at least one AE by system organ class and preferred term.

```{r, echo=FALSE, message=FALSE}
# use the summary data available in tfrmt
dat_ae <- tfrmt::data_ae |>
  dplyr::group_by(AEBODSYS, AETERM) |>
  dplyr::mutate(pct_high = value[col2 == "Xanomeline High Dose" & param == "pct"]) |>
  dplyr::ungroup() |>
  dplyr::filter(pct_high > 10) |>
  dplyr::select(-pct_high, -col1)

dat_ae_ptcpt <- dat_ae |>
  dplyr::filter(param %in% c("n", "pct"))

# create formatted display
tfrmt_ae <- tfrmt::tfrmt(
  group = AEBODSYS,
  label = AETERM,
  column = col2,
  param = param,
  value = value,
  sorting_cols = c(ord1, ord2),
  # specify value formatting
  body_plan = tfrmt::body_plan(
    tfrmt::frmt_structure(
      group_val = ".default", label_val = ".default",
      tfrmt::frmt_combine("{n} {pct}",
        n = tfrmt::frmt("XXX"),
        pct = tfrmt::frmt_when(
          "==100" ~ "",
          "==0" ~ "",
          TRUE ~ tfrmt::frmt("(xx%)")
        )
      )
    )
  ),
  col_plan = tfrmt::col_plan(
    -starts_with("ord")
  )
)

tfrmt::print_to_gt(tfrmt_ae, dat_ae_ptcpt) |>
  gt::tab_options(
    table.font.size = 12
  ) |>
  gt::cols_align(align = "center")
```

For the computations below, we will not only make use of the `ADAE` dataset, but we will also rely on `ADSL` for the full study population. To match between them, we need to do a small data manipulation on the naming of the treatment variable. 

```{r}
# rename trt variable
adsl <- ADSL |>
  dplyr::rename("TRTA" = "TRT01A")

# subset to Treatment emergent AES
adae <- ADAE |>
  dplyr::filter(TRTEMFL=="Y")
```

To get these participant level summaries, we will take a multi-step approach by calculating number and percent of subjects with: (1) at least one AE, (2) at least one AE by system organ class, and (3) at least one AE by system organ class and preferred term.

Because `ADAE` potentially contains multiple records per subject for a given AE, we must collapse the data so that it is subject-level. For the first step, we must reduce the data down to 1 record per subject with any AE (note that your data may contain derived flags to assist with this step). If it doesn't exist, we can create a flag variable to compute over:

```{r}
adae_subj <- adae |>
  dplyr::select(USUBJID, TRTA) |>
  dplyr::mutate(any_ae = 1) |>
  unique()
```

We then pass this dataset to `ard_categorical()` to compute the number and percentage of subjects by treatment arm. To ensure our denominator is the full study population, we pass our `adsl` data to the `denominator` argument:

```{r}
# any ae
ard_subj_any <- ard_categorical(
  data = adae_subj,
  by = TRTA,
  variables = any_ae,
  statistic = ~ c("n", "p"),
  denominator = adsl
)
```

For the second step, we create a new dataset of 1 record per subject with any AE by system organ class: 

```{r}
adae_soc <- adae |>
  dplyr::select(USUBJID, TRTA, AEBODSYS) |>
  unique()
```

We then pass this dataset to `ard_categorical()` to compute the number and percentage of subjects within each system organ class by treatment arm. 

```{r}
ard_subj_soc <- ard_categorical(
  data = adae_soc,
  by = TRTA,
  variables = AEBODSYS,
  statistic = ~ c("n", "p"),
  denominator = adsl
)
```

Finally, we create another dataset of 1 record per subject with any AE by system organ class and preferred term. 

```{r}
adae_pt <- adae |>
  dplyr::select(USUBJID, TRTA, AEBODSYS, AETERM) |>
  unique()
```

This time, we will use `ard_hierarchical()` to compute the number and percentage of subjects within each preferred term observed within each system organ class by treatment arm. The use of `ard_hierarchical()` ensures we keep the nested structure of preferred terms within system organ classes. 
 
```{r}
ard_subj_pt <- ard_hierarchical(
  data = adae_pt,
  by = TRTA,
  variables = c(AEBODSYS, AETERM),
  statistic = ~ c("n", "p"),
  denominator = adsl
)
```

All participant level summaries can be combined using `bind_ard()`.

```{r}
bind_ard(ard_subj_any, ard_subj_soc, ard_subj_pt)
```


#### Event-level Summaries

In addition to participant-level summaries, event-level summaries are often needed. For these types of tables, we report total counts of AEs , and therefore we can use the `ADAE` data directly. We will need to count AEs overall, by system organ class, and by preferred term (within system organ class).


```{r, echo=FALSE, message=FALSE}
dat_ae_event <- dat_ae |>
  dplyr::filter(param == "AEs")

# create formatted display
tfrmt_ae <- tfrmt::tfrmt(
  group = AEBODSYS,
  label = AETERM,
  column = col2,
  param = param,
  value = value,
  sorting_cols = c(ord1, ord2),
  # specify value formatting
  body_plan = tfrmt::body_plan(
    tfrmt::frmt_structure(group_val = ".default", label_val = ".default", frmt("xxx"))
  ),
  col_plan = tfrmt::col_plan(
    -starts_with("ord")
  )
)

tfrmt::print_to_gt(tfrmt_ae, dat_ae_event) |>
  gt::tab_options(
    table.font.size = 12
  ) |>
  gt::cols_align(align = "center")
```


For overall AE counts by treatment group, we can use `ard_hierarchical_count()` as such. As we don't have a hierarchy, this is equivalent to `ard_categorical()` with `statistic = ~ "n"`. We can rename the counts so they are distinguishable when stacked with the participate level counts. 
```{r}
# overall
ard_event_any <- ard_hierarchical_count(
  data = adae,
  variables = TRTA
)
```

For AE counts within preferred term and system organ class by treatment group, we can use `ard_hierarchical()` specifying our nesting as such:

```{r}
ard_event_socpt <- ard_hierarchical_count(
  data = adae,
  by = TRTA,
  variables = c(AEBODSYS, AETERM)
)
```
 
These event-level summaries can be combined using `bind_ard()`.
```{r}
bind_ard(ard_event_any, ard_event_socpt)
```
