---
title: "Updating ARD Objects"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
library(cards)
library(dplyr)
```

## Introduction

ARD (Analysis Results Data) objects are data frames that contain statistical summaries. Because they are data frames, any value in the ARD can be updated using standard dplyr or base R functions.

### Working with List Columns

ARD objects contain **list columns** such as `stat`, `fmt_fun`, `stat_label`, and others. List columns are columns where each cell can hold any R object—a number, a function, a vector, or even another data frame. If you're unfamiliar with list columns, they may look unusual when printed:

```{r}
ard <- ard_summary(ADSL, variables = AGE) |>
  select(variable, stat_name, stat, fmt_fun) |>
  head(3)
print(ard)
```

While list columns can be updated just like any other column in a data frame, the syntax can be less intuitive. For example, to update the `fmt_fun` column for specific statistics, you might write:
```{r}
ard |>
  mutate(
    fmt_fun = ifelse(stat_name %in% c("mean", "sd"), list(2L), fmt_fun)
  )
```

This requires wrapping values in `list()` and can become cumbersome for more complex updates.

### Helper Functions

To simplify working with these list columns, {cards} provides helper functions:

-   **`update_ard_fmt_fun()`**: Update formatting functions for specific statistics
-   **`update_ard_stat_label()`**: Update statistic labels

These functions handle the list column mechanics for you, making it easier to customize your ARD objects.

## Basic Usage

### Updating Formatting Functions

By default, statistics may use simple formatting. You can update the formatting function for specific statistics:

```{r}
# Create a basic ARD
ard <- ard_summary(ADSL, variables = AGE)

# Update formatting for mean and sd to show more decimal places
ard_updated <- ard |>
  update_ard_fmt_fun(
    stat_names = c("mean", "sd"),
    fmt_fun = 2L # 2 decimal places
  ) |>
  apply_fmt_fun()

# View results
ard_updated |>
  select(stat_name, stat, stat_fmt)
```

### Updating Statistic Labels

Combine formatting updates with custom labels:

```{r}
ard_summary(ADSL, variables = AGE) |>
  update_ard_fmt_fun(stat_names = c("mean", "sd"), fmt_fun = 1L) |>
  update_ard_stat_label(
    stat_names = c("mean", "sd"),
    stat_label = "Mean (SD)"
  ) |>
  apply_fmt_fun()
```

## Selective Updates

### Filtering by Variable

Update formatting for specific variables only:

```{r}
ard_summary(ADSL, variables = c(AGE, BMIBL)) |>
  update_ard_fmt_fun(
    variables = AGE, # Only update AGE
    stat_names = "mean",
    fmt_fun = 3L
  ) |>
  apply_fmt_fun()
```

### Filtering by Group

When working with stratified analyses, use the `filter` argument to target specific groups:

```{r}
# Update formatting only for the Placebo arm
ard_summary(
  ADSL,
  by = ARM,
  variables = AGE,
  statistic = ~ continuous_summary_fns(c("N", "mean"))
) |>
  update_ard_fmt_fun(
    stat_names = "mean",
    fmt_fun = 3L,
    filter = group1_level == "Placebo"
  ) |>
  apply_fmt_fun()
```

## Custom Formatting Functions

Beyond integer aliases, you can pass custom functions:

```{r}
# Custom formatter that adds parentheses
format_with_parens <- function(x) {
  paste0("(", format(round(x, 1), nsmall = 1), ")")
}

ard_summary(ADSL, variables = AGE) |>
  update_ard_fmt_fun(
    stat_names = "sd",
    fmt_fun = format_with_parens
  ) |>
  apply_fmt_fun()
```